---
title: "scWGBS - Lnuclei TSS"
output: html_notebook
---

```{r}
library(Rtsne)
library(tidyr)
library(dplyr)
library(ggplot2)
library(rlist)
library(pheatmap)
library(patchwork)
```

- 273 cells in original data 
- 273 cells passed filtering with Epiclomal 
- 897 non-redundant regions used for clustering by Epiclomal 
- Lnuclei split into two clusters
- Epiclomal Basic and Region result in same clusters 
- No batch effects based on pool 
- Results of Epiclomal shown both with t-SNE based on mean in each region and individual CpG sites 

## Epiclomal
- Selects regions with at leas 5% covrage in 10% of the cells
- Removes redundat regions by checking pearsons correlation between regions (0.95 cutoff used). The region with more missing data is removed. 
- CpG sites with no data accross all cells and cells with no data across CpG sites are removed.
- EpiclomalBasic: Assumens true hidden methylation states have same distribution accross all CpG sites  
- EpiclomalRegion: Allows for variation in distribution across regions 

Clustering: Non-probebalistic clustering methods used to initialize cluster assignment for Epiclomal  

- EucledianClust and DensityCut uses mean methylation vaues in each region to cluster cells
- HammingClust and PearsonClust uses individual methyation values at CpG sites to cluster cells 
- Results from the 4 non-probebalistic methods used as initial values in Epiclomal Basic and Region

Epiclomal uses bayesian  algorithm to infer both custer assignment and true hidden methylation states for each cluster. Chooses the best K and cluster assignmnt based on which combination of the two that produces the lowest DIC. 

## DATA 

```{r}
data_mean_region = read.csv("final_mean_meth_region_Lnuclei_TSS.tsv", sep = "\t")
data_cpg_sites = read.csv("input_Epiclomal_Lnuclei_TSS.tsv", sep = "\t")
cluster_region = read.csv("cluster_posteriors_bestrun_region_Lnuclei_TSS.tsv", sep = "\t")
keep_regions = read.csv("filtered_regions_Lnuclei_TSS.tsv", header = F)
```

```{r}
colnames(cluster_region)[2:3] = 1:2
cluster_region$cluster = colnames(cluster_region[2:3])[apply(cluster_region[2:3],1,which.max)]

cluster_region$col_cluster = ""
cluster_region[cluster_region$cluster == "1",]$col_cluster = "orange"
cluster_region[cluster_region$cluster == "2",]$col_cluster = "green4"
```

```{r}
data_mean_region[is.na(data_mean_region)] = 0
data_mean_region = as.data.frame(t(data_mean_region))
rownames(data_mean_region) = gsub("[.]", "-", rownames(data_mean_region))
data_mean_region = data_mean_region[,colnames(data_mean_region) %in% keep_regions$V1]
data_mean_region = data_mean_region[rownames(data_mean_region) %in% cluster_region$cell_id, ]
```

```{r}
data_cpg_sites[is.na(data_cpg_sites)] = 0
rownames(data_cpg_sites) = data_cpg_sites$cell_id
data_cpg_sites = data_cpg_sites[,!colnames(data_cpg_sites) %in% c("cell_id")]
```

### t-SNE Mean Region

```{r}
tsne_region <- Rtsne(data_mean_region, perplexity=20, check_duplicates = FALSE)
saveRDS(tsne_region, "tsne_region_Lnuclei_TSS.rds")

tsne_region = readRDS("tsne_region_Lnuclei_TSS.rds")
```

# PLOT

```{r}
tsne_plot_region = data.frame(x = tsne_region$Y[,1], y = tsne_region$Y[,2], cell_id = rownames(data_mean_region))
tsne_plot_region = merge(tsne_plot_region, cluster_region[c("cell_id", "cluster")], by = "cell_id")
```

### t-SNE plot by cluster 

```{r}
ggplot(tsne_plot_region, aes()) + 
  geom_point(aes(x=x, y=y, color=cluster), size = 2.5) +
  scale_colour_manual(values = c("orange", "green4", "royalblue3", "red3")) +
  xlab("dim 1") +
  ylab("dim 2") +
  ggtitle("TSS") +
  theme_bw()
```

### Heatmap by cluster 

```{r}
palette <- colorRampPalette(c("white", "blue"))(n = 299)
matrix = as.matrix(data_mean_region[cluster_region[order(cluster_region$cluster),]$cell_id,])

heatmap(matrix, col = palette, Rowv = NA, RowSideColors = cluster_region[order(cluster_region$cluster),]$col_cluster, xlab = "regions", ylab = "cells", labRow = NA, labCol = NA, scale = "col") 
legend(x="right", title = "cluster", legend = c("1", "2"), fill = c("orange", "green4")) 
```

### t-SNE CpG Sites 

```{r}
tsne_basic <- Rtsne(data_cpg_sites, perplexity=20, check_duplicates = FALSE)
saveRDS(tsne_basic, "tsne_basic_Lnuclei.rds")

tsne_basic = readRDS("tsne_basic_Lnuclei.rds")
```

# PLOT 

```{r}
tsne_plot_basic = data.frame(x = tsne_basic$Y[,1], y = tsne_basic$Y[,2], cell_id = rownames(data_cpg_sites))
tsne_plot_basic = merge(tsne_plot_basic, cluster_region[c("cell_id", "cluster")], by = "cell_id")
```

### t-SNE plot by cluster 

```{r}
ggplot(tsne_plot_basic, aes()) + 
  geom_point(aes(x=x, y=y, color=cluster), size = 2.5) +
  scale_colour_manual(values = c("orange", "green4")) +
  xlab("dim 1") +
  ylab("dim 2") +
  ggtitle("TSS") +
  theme_bw()
```

# t-SNE plot by pool

```{r}
tsne_plot_region$pool = unlist(strsplit(tsne_plot_region$cell_id, split = '-') %>% list.map(.[2]))
tsne_plot_basic$pool = unlist(strsplit(tsne_plot_basic$cell_id, split = '-') %>% list.map(.[2]))
```

```{r}
reg = ggplot(tsne_plot_region, aes()) + 
  geom_point(aes(x=x, y=y, color=pool), size = 2.5, show.legend = F) +
  xlab("dim 1") +
  ylab("dim 2") + 
  ggtitle("Mean Region") + 
  theme_bw()

bas = ggplot(tsne_plot_basic, aes()) + 
  geom_point(aes(x=x, y=y, color=pool), size = 2.5, show.legend = F) +
  xlab("dim 1") +
  ylab("dim 2") +
  ggtitle("CpG Sites") +
  theme_bw()

reg + bas
```

## t-SNE plot batch 

```{r}
tsne_plot_region$batch = "batch1"
tsne_plot_region[tsne_plot_region$pool %in% c("pool76", "pool77", "pool78", "pool79", "pool80", "pool81"),]$batch = "batch2"
tsne_plot_region[tsne_plot_region$pool %in% c("pool82", "pool83", "pool84", "pool85", "pool86", "pool87"),]$batch = "batch3"

tsne_plot_basic$batch = "batch1"
tsne_plot_basic[tsne_plot_basic$pool %in% c("pool76", "pool77", "pool78", "pool79", "pool80", "pool81"),]$batch = "batch2"
tsne_plot_basic[tsne_plot_basic$pool %in% c("pool82", "pool83", "pool84", "pool85", "pool86", "pool87"),]$batch = "batch3"
```

```{r}
reg = ggplot(tsne_plot_region, aes()) + 
  geom_point(aes(x=x, y=y, color=batch, shape = batch), size = 2.5, alpha = 0.7, show.legend = F) +
  scale_color_manual(values = c("green4", "red3", "royalblue3")) +
  xlab("dim 1") +
  ylab("dim 2") +
  ggtitle("Mean Region") + 
  theme_bw()

bas = ggplot(tsne_plot_basic, aes()) + 
  geom_point(aes(x=x, y=y, color=batch, shape = batch), size = 2.5, alpha = 0.7) +
  scale_color_manual(values = c("green4", "red3", "royalblue3")) +
  xlab("dim 1") +
  ylab("dim 2") +
  ggtitle("CpG Sites") + 
  theme_bw()

reg + bas
```

