---
title: "scWGBS - K562"
output: html_notebook
---

```{r}
library(Rtsne)
library(tidyr)
library(dplyr)
library(ggplot2)
library(rlist)
library(patchwork)
library("RColorBrewer")
```

- 230 cells in original data 
- 230 cells passed filtering with Epiclomal 
- 203 non-redundant regions used for clustering by Epiclomal 
- K562 one cluster 
- Epiclomal Basic and Region result in same clusters (Only EpiclomalRegion results shown here)
- Pool 34 & 35 does not cluster together 
- Batch effects can be seen 

## DATA 

```{r}
data_region = read.csv("final_mean_meth_region_K562.tsv", sep = "\t")
data_cpg_site = read.csv("input_Epiclomal_K562.tsv", sep = "\t")
cluster_region = read.csv("cluster_posteriors_bestrun_region_K562.tsv", sep = "\t")
```

```{r}
colnames(cluster_region)[2] = 1
cluster_region$cluster = colnames(cluster_region[2])[apply(cluster_region[2],1,which.max)]
cluster_region$col_cluster = c("orange")
```

```{r}
data_region[is.na(data_region)] = 0
data_region = as.data.frame(t(data_region))
rownames(data_region) = gsub("[.]", "-", rownames(data_region))
data_region = data_region[rownames(data_region) %in% cluster_region$cell_id,]
```

```{r}
data_cpg_site[is.na(data_cpg_site)] = 0
rownames(data_cpg_site) = data_cpg_site$cell_id
data_cpg_site = data_cpg_site[,!colnames(data_cpg_site) %in% c("cell_id")]
```

### t-SNE 

```{r}
#tsne_region = Rtsne(data_region, perplexity=30, check_duplicates = FALSE)
#saveRDS(tsne_region, "tsne_region.rds")

tsne_region = readRDS("tsne_region.rds")
```

```{r}
tsne_cpg_site = Rtsne(data_cpg_site, perplexity=30, check_duplicates = FALSE)
saveRDS(tsne_cpg_site, "tsne_cpg_site.rds")
```


# PLOT 

```{r}
tsne_plot_region = data.frame(x = tsne_region$Y[,1], y = tsne_region$Y[,2], cell_id = rownames(data_region))
tsne_plot_region = merge(tsne_plot_region, cluster_region[c("cell_id", "cluster")], by = "cell_id")
```

### t-SNE plot by cluster 

```{r}
ggplot(tsne_plot_region, aes()) + 
  geom_point(aes(x=x, y=y, color=cluster), size = 2.5) +
  scale_colour_manual(values = c("orange")) +
  xlab("dim 1") +
  ylab("dim 2") +
  theme_bw() +
  ggtitle("B") +
   theme(plot.title=element_text(size=30), 
         legend.text=element_text(size = 17),
         legend.title=element_text(size = 20))
```

### Heatmap by cluster 

```{r}
palette <- colorRampPalette(c("white", "blue"))(n = 299)
matrix = as.matrix(data_region[cluster_region[order(cluster_region$cluster),]$cell_id,])

heatmap(matrix, 
        col = palette, 
        Rowv = NA, 
        RowSideColors = cluster_region[order(cluster_region$cluster),]$col_cluster, 
        xlab = "regions", ylab = "cells", 
        labRow = NA, labCol = NA)
legend(x="right", legend = c("1"), fill = c("orange")) 
```

```{r}
tsne_plot_cpg_site = data.frame(x = tsne_cpg_site$Y[,1], y = tsne_cpg_site$Y[,2], cell_id = rownames(data_cpg_site))
tsne_plot_cpg_site = merge(tsne_plot_cpg_site, cluster_region[c("cell_id", "cluster")], by = "cell_id")
```

### t-SNE plot by cluster 

```{r}
ggplot(tsne_plot_cpg_site, aes()) + 
  geom_point(aes(x=x, y=y, color=cluster), size = 2.5) +
  scale_colour_manual(values = c("orange")) +
  xlab("dim 1") +
  ylab("dim 2") +
  ggtitle("CpG Islands") +
  theme_bw()
```

# t-SNE plot by pool

```{r}
tsne_plot_region$pool = unlist(strsplit(tsne_plot_region$cell_id, split = '-') %>% list.map(.[2]))

tsne_plot_cpg_site$pool = unlist(strsplit(tsne_plot_cpg_site$cell_id, split = '-') %>% list.map(.[2]))
```

```{r}
reg = ggplot(tsne_plot_region, aes()) + 
  geom_point(aes(x=x, y=y, color=pool), size = 2.5, show.legend = F) +
  xlab("dim 1") +
  ylab("dim 2") +
  ggtitle("Mean Reagion") +
  theme_bw()

cpg = ggplot(tsne_plot_cpg_site, aes()) + 
  geom_point(aes(x=x, y=y, color=pool), size = 2.5) +
  xlab("dim 1") +
  ylab("dim 2") +
  ggtitle("CpG Sites") +
  theme_bw()
reg + cpg

```

### t-SNE plot pool 34 & 35 

```{r}
pool3435_reg = tsne_plot_region %>% 
           filter(pool == "pool34" | pool == "pool35")

pool3435_cpg = tsne_plot_cpg_site %>% 
           filter(pool == "pool34" | pool == "pool35")
```

```{r}
reg = ggplot(tsne_plot_region, aes()) + 
  geom_point(aes(x=x, y=y, color=cluster), size = 1.5, color = "grey") +
  xlab("dim 1") +
  ylab("dim 2") +
  theme_bw() +
  geom_point(data = pool3435_reg, aes(x=x, y=y, shape = pool), size = 2.5, color = "red",  show.legend = F) 

cpg = ggplot(tsne_plot_cpg_site, aes()) + 
  geom_point(aes(x=x, y=y, color=cluster), size = 1.5, color = "grey") +
  xlab("dim 1") +
  ylab("dim 2") +
  theme_bw() +
  geom_point(data = pool3435_cpg, aes(x=x, y=y, shape = pool), size = 2.5, color = "red") 

reg + cpg 

```

## t-SNE plot batch 

```{r}
tsne_plot_region$Batch = "1"
tsne_plot_region[tsne_plot_region$pool %in% c("pool70", "pool71", "pool72", "pool73", "pool74", "pool75"),]$Batch = "2"
tsne_plot_region[tsne_plot_region$pool %in% c("pool34", "pool35"),]$Batch = "snmC-seq2"

tsne_plot_cpg_site$batch = "batch1"
tsne_plot_cpg_site[tsne_plot_cpg_site$pool %in% c("pool70", "pool71", "pool72", "pool73", "pool74", "pool75"),]$batch = "batch2"
tsne_plot_cpg_site[tsne_plot_cpg_site$pool %in% c("pool34", "pool35"),]$batch = "snmC-seq2"
```

```{r}
reg = ggplot(tsne_plot_region, aes()) + 
  geom_point(aes(x=x, y=y, color=batch, shape = batch), size = 2.5, alpha = 0.8,  show.legend = F) +
  scale_color_manual(values = c( "royalblue3", "green4", "red3")) +
  xlab("dim 1") +
  ylab("dim 2") +
  theme_bw()

cpg = ggplot(tsne_plot_cpg_site, aes()) + 
  geom_point(aes(x=x, y=y, color=batch, shape = batch), size = 2.5, alpha = 0.8) +
  scale_color_manual(values = c("red3", "royalblue3", "green4")) +
  xlab("dim 1") +
  ylab("dim 2") +
  theme_bw()

reg + cpg

```

```{r}
ggplot(tsne_plot_region, aes()) + 
  geom_point(aes(x=x, y=y, color=Batch, shape = Batch), size = 2.5, alpha = 0.8) +
  scale_color_manual(values = c("royalblue3", "green4", "red3")) +
  xlab("dim 1") +
  ylab("dim 2") +
  ggtitle("D") +
  theme_bw() +
  theme(plot.title=element_text(size=30), 
         legend.text=element_text(size = 17),
         legend.title=element_text(size = 20))
```

